---
title: "`fluxible`: an R package to process ecosystem gas fluxes from closed-loop chambers in an automated and reproducible way"
subtitle: "`fluxible`: ecosystem gas fluxes calculations"
# format: docx # tbl-packages need to be passed as comments for docx
format:
  pdf:
#  aog-article-pdf: 
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    number-sections: true
    number-depth: 4
    keep-tex: true
    colorlinks: false
    # linestretch: 2
keywords: ecosystem gas fluxes, flux chambers, R package, reproducible workflow
bibliography: /home/jga051/Dropbox/PhD/biblio_phd_zot.bib
csl: /home/jga051/Documents/01_PhD/emerald-harvard.csl
date: today
execute:
  echo: false
---

**Key-words:** ecosystem gas fluxes, flux chambers, greenhouse gas emissions, R package, reproducible workflow


```{r}
#| label: setup
#| include: false

library(googlesheets4)
library(plume)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(tidyverse.quiet = TRUE)

tbl_authors <- read_sheet(gs4_find("authors_fluxible"))

aut <- Plume$new(tbl_authors)
aut$set_corresponding_authors(1, 5)
```

`r aut$get_author_list("^a,^co")`

```{r}
#| results: asis
aut$get_affiliations() |> cat(sep = "\n\n")
```

\*Correspondence to: `r aut$get_contact_details()`

# Acknowledgements {.unnumbered}
The `fluxible` R package is the result of discussions and methodology conceptualization about ecosystem gas fluxes data processing during the Plant Functional Traits Courses (PFTC) 6, and the ThreeD and INCLINE projects.
The ThreeD project was funded by the Research Council of Norway KLIMAFORSK project 287801 (2019 – 2023).
The INCLINE project was funded by FRIMEDBIO project 274712 (2018 – 2022).
The PFTC6 course was funded under the Research Council of Norway projects RECITE INTPART project 274831 (2018 – 2021), ExperTS INTPART project 287784 (2019 – 2022), and the Centre for Internationalisation in Education (SIU) project TraitTrain HNP-2015/10037 (2016 – 2018).
We also would like to thank Vincent Belde who, as an early user of `fluxible` during his master thesis, came up with valuable improvement suggestions, and Adam Klimes for his help in analytics and R coding.

# Data availability {.unnumbered}
The data used in the validation (Section 3) and those used in the data preparation example (Appendix B) can be downloaded from Open Science Framework: [osf.io/srj24](https://osf.io/srj24/).
The data for the detailed example (Appendix A) and the workflow with two gases (Appendix C) are part of `fluxible`.
The code of the package is available on GitHub: [github.com/Plant-Functional-Trait-Course/fluxible](https://github.com/Plant-Functional-Trait-Course/fluxible).
The code for the validation section and figures production is available on GitHub: [github.com/jogaudard/fluxible-MEE](https://github.com/jogaudard/fluxible-MEE).

# Conflict of interest {.unnumbered}
The authors declare no conflict of interest.

# Authors contribution {.unnumbered}

All students and collaborators in the courses and projects where the conceptualization of the `fluxible` R package was discussed were offered to participate in the manuscript.
```{r}
#| results: asis
#| echo: false
aut$get_contributions(literal_names = TRUE, by_author = FALSE, roles_first = FALSE) |> cat(sep = "; ")
```
