---
title: "Code to produce the datasets used in the Fluxible manuscript"
bibliography: /home/jga051/Dropbox/PhD/biblio_phd_zot.bib
csl: /home/jga051/Documents/01_PhD/emerald-harvard.csl
date: today
format: pdf
author: Joseph Gaudard
---

Since LI-COR sent so much data and it is heavy to handle, I will prepare them in a separate document to avoid slowing down the document with the manuscript.

```{r}
#| echo: FALSE

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(tidyverse.quiet = TRUE)
```

# Data for validation

## Data import
```{r}
#| label: osf-import
#| message: true
#| warning: false
#| eval: true
library(dataDownloader)

get_file(
    node = "srj24",
    file = "validation_data.zip"
)

unzip("validation_data.zip")


# saving space
file.remove("validation_data.zip")
```


```{r}
#| label: licor-import
#| message: FALSE
#| warning: false
#| eval: true


library(tidyverse)

# some work needed to get around the triple header

data_name1 <- read_csv(
      "validation_data/licor_fluxes.csv",
      n_max = 1,
      col_names = FALSE,
      show_col_types = FALSE,
      progress = FALSE
    )

  data_name1 <- as.character(data_name1[1, ])

  data_name2 <- read_csv(
      "validation_data/licor_fluxes.csv",
      skip = 1,
      n_max = 1,
      col_names = FALSE,
      show_col_types = FALSE,
      progress = FALSE
    )

  data_name2 <- as.character(data_name2[1, ])

  data_name <- paste(
    data_name1,
    data_name2,
    sep = "_"
  )

licor_fluxes_og <- read_csv(
    "validation_data/licor_fluxes.csv",
    skip = 3,
    col_names = data_name
)


licor_fluxes <- licor_fluxes_og |>
    mutate(
        datetime = ymd_hms(`LI-8250_DATE_TIME initial_value`)
    ) |>
    select(!c(
        `LI-8250_DATE_TIME initial_value`,
        `LI-8250_PORT`,
        `8250-01_PORT`
        )) |>
    arrange(datetime) |>
    rowid_to_column("row_id") |>
    pivot_longer(
        !c(datetime, row_id),
        names_to = "method",
        values_to = "licor_flux"
    )

saveRDS(licor_fluxes, "validation_data/licor_fluxes.rds")
# we will use RDS format to save intermediate objects in case it crashes
rm(list = c(
    "licor_fluxes", "licor_fluxes_og", "data_name", "data_name1", "data_name2"
    ))
```

```{r}
#| label: raw-import
#| message: FALSE
#| warning: false
#| eval: true

library(licoread)

all_fluxes <- licoread(
    "validation_data/export"
    # sample = 10 # for testing with a smaller file
    )

# saveRDS in case it crashes
# no need to rerun everything
saveRDS(all_fluxes, "validation_data/licor_testing_fluxes.rds")

# saving space on disk now that we have rds files
unlink("validation_data/export", recursive = TRUE)
```

```{r}
#| label: functions
#| message: FALSE
#| warning: false

library(fluxible)
# writing a fluxible wrapper function to save time

flux_compare <- function(
    conc_df,
    fit_type,
    start_cut = 20,
    end_cut = 0,
    conc_unit,
    flux_unit,
    method,
    force_discard = c(),
    rsquared_threshold = 0.7
) {
        
    conc_fit <- conc_df |>
    flux_fitting(
        start_cut = start_cut,
        end_cut = end_cut,
        fit_type = fit_type
        )

    conc_flag <- conc_fit |>
    flux_quality(
        f_conc = f_conc,
        instr_error = 10,
        force_discard = force_discard,
        #just getting rid of the start error since we have different gases and units
        rsquared_threshold = rsquared_threshold,
        error = 100000)

    flux <- conc_flag |>
    flux_calc(
        slope_col = f_slope_corr,
        f_datetime = f_datetime,
        temp_air_col = CHAMBER_TA,
        setup_volume = chamber_vol,
        plot_area = plot_area,
        atm_pressure = 1,
        conc_unit = conc_unit,
        flux_unit = flux_unit,
        cols_keep = "f_quality_flag"
    ) |>
    mutate( # licor wants umol/m2/s
        method = method,
        f_flux = f_flux/3600
    )

    flux
}

```

## LI-7810_CO2_DRY

```{r}
#| label: import-co2_7810
#| message: FALSE
#| warning: FALSE
#| eval: true

# all_fluxes <- readRDS("validation_data/licor_testing_fluxes.rds")


list_gases(all_fluxes)

co2_7810 <- licoread_to_fluxible(
    all_fluxes,
    focus_gas = "LI-7810_CO2_DRY",
    datetime_col = c("LI-8250_DATE", "LI-8250_TIME")
)

co2_7810 <- co2_7810 |>
     select(
        f_fluxid,
        f_datetime,
        f_start,
        f_end,
        f_conc,
        CHAMBER_TB,
        CHAMBER_TA,
        CHAMBER_TS_1,
        "METADATA_VOLUME_TOTAL_cm+3",
        "CHAMBER_AREA_cm+2"
        ) |>
        mutate(
            chamber_vol = `METADATA_VOLUME_TOTAL_cm+3` * 0.001,
            plot_area = `CHAMBER_AREA_cm+2` * 0.0001
        ) |>
        select(
            !c(
                "METADATA_VOLUME_TOTAL_cm+3",
                "CHAMBER_AREA_cm+2"
            )
        )

saveRDS(co2_7810, "validation_data/co2_7810.rds")

```

```{r}
#| label: fluxible-co2_7810
#| message: FALSE
#| warning: FALSE
#| eval: true

# co2_7810 <- readRDS("validation_data/co2_7810.rds")
    
co2_7810_lm <- flux_compare(
    co2_7810,
    fit_type = "linear",
    conc_unit = "ppm",
    flux_unit = "micromol",
    method = "LI-7810_FCO2_DRY LIN"
)
saveRDS(co2_7810_lm, "validation_data/co2_7810_lm.rds")


co2_7810_hm <- flux_compare(
    co2_7810,
    fit_type = "exp_hm",
    conc_unit = "ppm",
    flux_unit = "micromol",
    method = "LI-7810_FCO2_DRY"
)
saveRDS(co2_7810_hm, "validation_data/co2_7810_hm.rds")
rm(co2_7810)


```

## LI-7810_CH4_DRY

```{r}
#| label: import-ch4_7810
#| message: FALSE
#| warning: FALSE
#| eval: true

# all_fluxes <- readRDS("validation_data/licor_testing_fluxes.rds")

ch4_7810 <- licoread_to_fluxible(
    all_fluxes,
    focus_gas = "LI-7810_CH4_DRY",
    datetime_col = c("LI-8250_DATE", "LI-8250_TIME")
)

ch4_7810 <- ch4_7810 |>
     select(
        f_fluxid,
        f_datetime,
        f_start,
        f_end,
        f_conc,
        CHAMBER_TB,
        CHAMBER_TA,
        CHAMBER_TS_1,
        "METADATA_VOLUME_TOTAL_cm+3",
        "CHAMBER_AREA_cm+2"
        ) |>
        mutate(
            chamber_vol = `METADATA_VOLUME_TOTAL_cm+3` * 0.001,
            plot_area = `CHAMBER_AREA_cm+2` * 0.0001
        ) |>
        select(
            !c(
                "METADATA_VOLUME_TOTAL_cm+3",
                "CHAMBER_AREA_cm+2"
            )
        )

saveRDS(ch4_7810, "validation_data/ch4_7810.rds")


```

```{r}
#| label: fluxible-ch4_7810
#| message: FALSE
#| warning: false
#| eval: true


# ch4_7810 <- readRDS("validation_data/ch4_7810.rds")

ch4_7810_lm <- flux_compare(
    ch4_7810,
    fit_type = "linear",
    conc_unit = "ppb",
    flux_unit = "micromol",
    method = "LI-7810_FCH4_DRY LIN"
) |>
    mutate(
        f_flux = f_flux * 1000 # SFP has it in nanomol/m2/s
    )
saveRDS(ch4_7810_lm, "validation_data/ch4_7810_lm.rds")

ch4_7810_hm <- flux_compare(
    ch4_7810,
    fit_type = "exp_hm",
    conc_unit = "ppb",
    flux_unit = "micromol",
    method = "LI-7810_FCH4_DRY"
) |>
    mutate(
        f_flux = f_flux * 1000 # SFP has it in nanomol/m2/s
    )
saveRDS(ch4_7810_hm, "validation_data/ch4_7810_hm.rds")
rm(ch4_7810)

```

## LI-7820_N2O_DRY

```{r}
#| label: import-n2o_7820
#| message: FALSE
#| warning: FALSE
#| eval: true

# all_fluxes <- readRDS("validation_data/licor_testing_fluxes.rds")

n2o_7820 <- licoread_to_fluxible(
    all_fluxes,
    focus_gas = "LI-7820_N2O_DRY",
    datetime_col = c("LI-8250_DATE", "LI-8250_TIME")
)

n2o_7820 <- n2o_7820 |>
     select(
        f_fluxid,
        f_datetime,
        f_start,
        f_end,
        f_conc,
        CHAMBER_TB,
        CHAMBER_TA,
        CHAMBER_TS_1,
        "METADATA_VOLUME_TOTAL_cm+3",
        "CHAMBER_AREA_cm+2"
        ) |>
        mutate(
            chamber_vol = `METADATA_VOLUME_TOTAL_cm+3` * 0.001,
            plot_area = `CHAMBER_AREA_cm+2` * 0.0001
        ) |>
        select(
            !c(
                "METADATA_VOLUME_TOTAL_cm+3",
                "CHAMBER_AREA_cm+2"
            )
        )

saveRDS(n2o_7820, "validation_data/n2o_7820.rds")

```

```{r}
#| label: fluxible-n2o_7820
#| message: FALSE
#| warning: false
#| eval: true

# n2o_7820 <- readRDS("validation_data/n2o_7820.rds")

n2o_7820_lm <- flux_compare(
    n2o_7820,
    fit_type = "linear",
    conc_unit = "ppb",
    flux_unit = "micromol",
    method = "LI-7820_FN2O_DRY LIN"
) |>
    mutate(
        f_flux = f_flux * 1000 # SFP has it in nanomol/m2/s
    )
saveRDS(n2o_7820_lm, "validation_data/n2o_7820_lm.rds")


n2o_7820_hm <- flux_compare(
    n2o_7820,
    fit_type = "exp_hm",
    conc_unit = "ppb",
    flux_unit = "micromol",
    method = "LI-7820_FN2O_DRY"
) |>
    mutate(
        f_flux = f_flux * 1000 # SFP has it in nanomol/m2/s
    )
saveRDS(n2o_7820_hm, "validation_data/n2o_7820_hm.rds")

rm(n2o_7820)

```

## LI-870_CO2_DRY

```{r}
#| label: import-co2_870
#| message: FALSE
#| warning: FALSE
#| eval: true

# all_fluxes <- readRDS("validation_data/licor_testing_fluxes.rds")

co2_870 <- licoread_to_fluxible(
    all_fluxes,
    focus_gas = "LI-870_CO2_DRY",
    datetime_col = c("LI-8250_DATE", "LI-8250_TIME")
)

co2_870 <- co2_870 |>
     select(
        f_fluxid,
        f_datetime,
        f_start,
        f_end,
        f_conc,
        CHAMBER_TB,
        CHAMBER_TA,
        CHAMBER_TS_1,
        "METADATA_VOLUME_TOTAL_cm+3",
        "CHAMBER_AREA_cm+2"
        ) |>
        mutate(
            chamber_vol = `METADATA_VOLUME_TOTAL_cm+3` * 0.001,
            plot_area = `CHAMBER_AREA_cm+2` * 0.0001
        ) |>
        select(
            !c(
                "METADATA_VOLUME_TOTAL_cm+3",
                "CHAMBER_AREA_cm+2"
            )
        )

saveRDS(co2_870, "validation_data/co2_870.rds")

rm(all_fluxes)

```

```{r}
#| label: fluxible-co2_870
#| message: FALSE
#| warning: false
#| eval: true

# co2_870 <- readRDS("validation_data/co2_870.rds")

co2_870_lm <- flux_compare(
    co2_870,
    fit_type = "linear",
    conc_unit = "ppm",
    flux_unit = "micromol",
    method = "LI-870_FCO2_DRY LIN"
)
saveRDS(co2_870_lm, "validation_data/co2_870_lm.rds")


co2_870_hm <- flux_compare(
    co2_870,
    fit_type = "exp_hm",
    conc_unit = "ppm",
    flux_unit = "micromol",
    method = "LI-870_FCO2_DRY"
)
saveRDS(co2_870_hm, "validation_data/co2_870_hm.rds")
rm(co2_870)

```


## Gathering

```{r}
#| label: gathering-sfp
#| message: FALSE
#| warning: false
#| eval: true

co2_7810_hm <- readRDS("validation_data/co2_7810_hm.rds")
co2_7810_lm <- readRDS("validation_data/co2_7810_lm.rds")
ch4_7810_hm <- readRDS("validation_data/ch4_7810_hm.rds")
ch4_7810_lm <- readRDS("validation_data/ch4_7810_lm.rds")
n2o_7820_hm <- readRDS("validation_data/n2o_7820_hm.rds")
n2o_7820_lm <- readRDS("validation_data/n2o_7820_lm.rds")
co2_870_hm <- readRDS("validation_data/co2_870_hm.rds")
co2_870_lm <- readRDS("validation_data/co2_870_lm.rds")


fluxible_fluxes_hm20 <- bind_rows(
    co2_7810_hm,
    co2_7810_lm,
    ch4_7810_hm,
    ch4_7810_lm,
    n2o_7820_hm,
    n2o_7820_lm,
    co2_870_hm,
    co2_870_lm
) |>
arrange(f_datetime) |>
mutate(
    .by = f_fluxid,
    row_id = cur_group_id()
)
saveRDS(fluxible_fluxes_hm20, "validation_data/fluxible_fluxes_hm20.rds")

licor_fluxes <- readRDS("validation_data/licor_fluxes.rds")
# fluxible_fluxes_hm20 <- readRDS("validation_data/fluxible_fluxes_hm20.rds")
rm(list = c(
    "co2_7810_hm",
    "co2_7810_lm",
    "ch4_7810_hm",
    "ch4_7810_lm",
    "n2o_7820_hm",
    "n2o_7820_lm",
    "co2_870_hm",
    "co2_870_lm"
    ))

comparison_fluxes_hm20 <- left_join(
    licor_fluxes,
    fluxible_fluxes_hm20,
    by = join_by(row_id, method)
    ) |>
    # drop_na(f_flux) |>
    mutate(
        gas = str_extract(method, "\\w*[-]\\w*"),
        model = case_when(
            f_model == "exp_hm" ~ "Exponential",
            f_model == "linear" ~ "Linear"
        )
    )

saveRDS(comparison_fluxes_hm20, "validation_data/comparison_fluxes_hm20.rds")
rm(list = c("licor_fluxes", "fluxible_fluxes_hm20"))
```

## Figure
```{r}
#| label: fig-sfp
#| echo: false
#| message: FALSE
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| eval: true
#| fig-cap: "Comparison of fluxes output between SoilFluxPro (SFP; y-axis) and the Fluxible R package (x-axis) for CO~2~, CH~4~ and N~2~O, measured with an LI-8250 soil chamber system using instruments LI-870, LI-7820, and LI-7810, with 8 8200-104 chambers. Continuous lines show the 1:1 line. Both linear (purple) and exponential (green) models were used while in Fluxible, the `exp_hm` model was used as it is the closest to the SFP workflow. The first 20 s of each measurement were cut before fitting in Fluxible, which is conservatively matching the deadband used in SoilFluxPro. The CH~4~ data could not be fitted with the HM model because of the absence of curvature."

library(ggplot2)
library(ggpmisc)
library(viridis)



# comparison_fluxes_hm20 <- readRDS("validation_data/comparison_fluxes_hm20.rds")

# flux_flag_count(comparison_fluxes_hm20)

gas_names <- c(
  `LI-7810_FCH4_DRY` = "a) LI-7810 CH4",
  `LI-7810_FCO2_DRY` = "b) LI-7810 CO2",
  `LI-7820_FN2O_DRY` = "c) LI-7820 N2O",
  `LI-870_FCO2_DRY` = "d) LI-870 CO2"
)

plot_sfp_hm <- comparison_fluxes_hm20 |>
    filter(
      method != "LI-7810_FCH4_DRY"
    ) |>
    ggplot(aes(f_flux, licor_flux, color = model)) +
    theme_bw() +
    scale_color_manual(values = c(
        Exponential = "#0c9f0b",
        Linear = "#bf28bd")
    ) +
    # theme(legend.position = c(0.87, 0.25)) +
    # scale_fill_viridis(discrete = TRUE) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    stat_poly_line() +
    # stat_correlation(use_label("cor.label", "R2", "n")) +
    stat_poly_eq(use_label("R2", "n")) +
    facet_wrap(gas ~ ., scales = "free", nrow = 2, labeller = as_labeller(gas_names)) +
    labs(
        title = "Comparison of flux estimates between SoilFluxPro and Fluxible",
        x = "Fluxes calculated with Fluxible (exp_hm and linear model respectively)",
        y = "Fluxes as provided by LI-COR (calculated with SFP)",
        color = "Model"
    )
plot_sfp_hm
```

## Investigation of CH4

```{r}
#| label: ch4-investigation
#| echo: true
#| fig-width: 10
#| fig-height: 10
#| message: FALSE
#| warning: false


# comparison_fluxes_hm20 <- readRDS("validation_data/comparison_fluxes_hm20.rds")

investigation_df <- comparison_fluxes_hm20 |>
    filter(
        gas == "LI-7810_FCH4_DRY"
        & model == "Linear"
        & f_quality_flag == "zero"
    ) |>
    mutate(
        diff = abs(licor_flux - f_flux)
    ) |>
    slice_max(abs(licor_flux), n = 5)

prob_id <- pull(investigation_df, f_fluxid)


ch4_7810 <- readRDS("validation_data/ch4_7810.rds")


ch4_prob <- ch4_7810 |>
    filter(
        f_fluxid %in% prob_id
    )

rm(ch4_7810)

ch4_prob_slope <- flux_fitting(
    ch4_prob,
    start_cut = 5,
    fit_type = "linear",
    end_cut = 90
)

ch4_prob_flag <- flux_quality(
    ch4_prob_slope,
    instr_error = 10, # same parameters used before for consistency
    error = 100000
)

flux_plot(
    ch4_prob_flag,
    f_ylim_upper = 9150,
    f_ylim_lower = 3700,
    facet_wrap_args = list(nrow = 1, scales = "free")
)

# flux_plot(
#     ch4_prob_flag,
#     f_ylim_upper = 2500,
#     f_ylim_lower = 2000,
#     # facet_wrap_args = list(nrow = 1, scales = "free"),
#     print_plot = FALSE,
#     output = "pdfpages"
# )

```

It seems that some CH~4~ fluxes should have a different focus window.
If it was a normal field campaign data processing we could isolate them and apply a different cut.
But there are 16'000+ measurements, and the point is to compare the method anyway.

Let's try anyway:
```{r}
#| label: ch4-investigationii
#| message: FALSE
#| warning: false
#| fig-width: 10
#| fig-height: 8


# comparison_fluxes_hm20 <- readRDS("validation_data/comparison_fluxes_hm20.rds")

zero_df <- comparison_fluxes_hm20 |>
    filter(
        gas == "LI-7810_FCH4_DRY"
        & model == "Linear"
        & f_quality_flag == "zero"
    ) |>
    select(f_fluxid, licor_flux, f_datetime) |>
    arrange(f_datetime) |>
    rowid_to_column("row_id") |>
    select(!f_datetime)


zero_id <- pull(zero_df, f_fluxid)


ch4_7810 <- readRDS("validation_data/ch4_7810.rds")


ch4_zero <- ch4_7810 |>
    filter(
        f_fluxid %in% zero_id
    )


ch4_redo <- flux_compare(
    ch4_zero,
    "linear",
    start_cut = 40,
    end_cut = 0,
    conc_unit = "ppb",
    flux_unit = "micromol",
    method = "LI-7810_FCH4_DRY LIN",
    rsquared_threshold = 0.01, # filtering with R2 square is not appropriate since SFP does not flag zero fluxes
    force_discard = c(
        "82m-0106-20210304073410", # positive in the second half
                                   # focus window does not match
        "82m-0106-20210326005440" # lots of noise at the start
                                  # focus window does not match?
    )
    ) |>
    mutate(
        f_flux = f_flux * 1000 # SFP has it in nanomol/m2/s
        )

flux_flag_count(ch4_redo)

ch4_test <- ch4_redo |>
    select(f_fluxid, f_flux, f_datetime) |>
    arrange(f_datetime) |>
    rowid_to_column("row_id") |>
    left_join(zero_df, by = c("row_id", "f_fluxid"))

ch4_test |>
    ggplot(aes(f_flux, licor_flux)) +
    theme_bw() +
    geom_point() +
    geom_abline(slope = 1, intercept = 0)

# ch4_test_prob <- ch4_test |>
#     mutate(
#         diff = abs(licor_flux - f_flux)
#     ) |>
#     slice_max(diff, n = 1)
#     # slice_max(f_flux, n = 1)

# prob_id <- pull(ch4_test_prob, f_fluxid)

# ch4_7810 <- readRDS("validation_data/ch4_7810.rds")

# ch4_prob <- ch4_7810 |>
#     filter(
#         f_fluxid %in% prob_id
#     )

# rm(ch4_7810)

# ch4_prob_slope <- flux_fitting(
#     ch4_prob,
#     start_cut = 40,
#     fit_type = "linear",
#     end_cut = 0
# )

# ch4_prob_flag <- flux_quality(
#     ch4_prob_slope,
#     instr_error = 10, # same parameters used before for consistency
#     error = 100000
# )

# flux_plot(
#     ch4_prob_flag,
#     f_ylim_upper = 16000,
#     f_ylim_lower = 15800,
#     facet_wrap_args = list(nrow = 1, scales = "free")
# )

ch4_7810_lm <- readRDS("validation_data/ch4_7810_lm.rds")

ch4_ok <- ch4_7810_lm |>
    filter(
        !(f_fluxid %in% zero_id)
    )

flux_flag_count(ch4_ok)

ch4_7810_lm <- bind_rows(ch4_ok, ch4_redo)

flux_flag_count(ch4_7810_lm)

saveRDS(ch4_7810_lm, "validation_data/ch4_7810_lmii.rds")

rm(ch4_7810)
```


## Investigation of CO2 lin 870
```{r}
#| label: co2-investigation
#| echo: true
#| fig-width: 10
#| fig-height: 10

investigation_df <- comparison_fluxes_hm20 |>
    filter(
        gas == "LI-870_FCO2_DRY"
        & model == "Linear"
    ) |>
    mutate(
        diff = abs(licor_flux - f_flux)
    ) |>
    slice_max(diff, n = 1)

prob_id <- pull(investigation_df, f_fluxid)


co2_870 <- readRDS("validation_data/co2_870.rds")


co2_prob <- co2_870 |>
    filter(
        f_fluxid %in% prob_id
    )

rm(co2_870)

co2_prob_slope <- flux_fitting(
    co2_prob,
    start_cut = 20,
    fit_type = "linear",
    end_cut = 0
)

co2_prob_flag <- flux_quality(
    co2_prob_slope,
    instr_error = 10, # same parameters used before for consistency
    error = 100000
)


flux_plot(
    co2_prob_flag,
    f_ylim_upper = 5000,
    f_ylim_lower = -10000
)

```

## Reprocessing after investigation


```{r}
#| label: fluxible-co2_870ii
#| message: FALSE
#| warning: false

co2_870 <- readRDS("validation_data/co2_870.rds")

co2_870_lm <- flux_compare(
    co2_870,
    fit_type = "linear",
    conc_unit = "ppm",
    flux_unit = "micromol",
    method = "LI-870_FCO2_DRY LIN",
    force_discard = "82m-0106-20210419053045" # has -9999 part of the flux
                                              # probably instrument disconnected
)
saveRDS(co2_870_lm, "validation_data/co2_870_lmii.rds")

rm(co2_870)

```


## Gathering II

```{r}
#| label: gathering-sfpii
#| message: FALSE
#| warning: false
#| eval: true

co2_7810_hm <- readRDS("validation_data/co2_7810_hm.rds")
co2_7810_lm <- readRDS("validation_data/co2_7810_lm.rds")
ch4_7810_hm <- readRDS("validation_data/ch4_7810_hm.rds")
ch4_7810_lm <- readRDS("validation_data/ch4_7810_lmii.rds")
n2o_7820_hm <- readRDS("validation_data/n2o_7820_hm.rds")
n2o_7820_lm <- readRDS("validation_data/n2o_7820_lm.rds")
co2_870_hm <- readRDS("validation_data/co2_870_hm.rds")
co2_870_lm <- readRDS("validation_data/co2_870_lmii.rds")
# co2_870_lm <- readRDS("validation_data/co2_870_lm.rds")


fluxible_fluxes_hm20 <- bind_rows(
    co2_7810_hm,
    co2_7810_lm,
    ch4_7810_hm,
    ch4_7810_lm,
    n2o_7820_hm,
    n2o_7820_lm,
    co2_870_hm,
    co2_870_lm
) |>
arrange(f_datetime) |>
mutate(
    .by = f_fluxid,
    row_id = cur_group_id()
)
saveRDS(fluxible_fluxes_hm20, "validation_data/fluxible_fluxes_hm20ii.rds")

licor_fluxes <- readRDS("validation_data/licor_fluxes.rds")
# fluxible_fluxes_hm20 <- readRDS("validation_data/fluxible_fluxes_hm20.rds")
rm(list = c(
    "co2_7810_hm",
    "co2_7810_lm",
    "ch4_7810_hm",
    "ch4_7810_lm",
    "n2o_7820_hm",
    "n2o_7820_lm",
    "co2_870_hm",
    "co2_870_lm"
    ))

comparison_fluxes_hm20 <- left_join(
    licor_fluxes,
    fluxible_fluxes_hm20,
    by = join_by(row_id, method)
    ) |>
    # drop_na(f_flux) |>
    mutate(
        gas = str_extract(method, "\\w*[-]\\w*"),
        model = case_when(
            f_model == "exp_hm" ~ "Exponential",
            f_model == "linear" ~ "Linear"
        )
    )

saveRDS(comparison_fluxes_hm20, "comparison_fluxes_hm20ii.rds")
rm(list = c("licor_fluxes", "fluxible_fluxes_hm20"))
```

## Figure II
```{r}
#| label: fig-sfpii
#| echo: false
#| message: FALSE
#| warning: false
#| fig-width: 10
#| fig-height: 8
#| fig-cap: "Comparison of fluxes output between SoilFluxPro (SFP; y-axis) and the Fluxible R package (x-axis) for CO~2~, CH~4~ and N~2~O, measured with an LI-8250 soil chamber system using instruments LI-870, LI-7820, and LI-7810, with 8 8200-104 chambers. Continuous lines show the 1:1 line. Both linear (purple) and exponential (green) models were used while in Fluxible, the `exp_hm` model was used as it is the closest to the SFP workflow. The first 20 s of each measurement were cut before fitting in Fluxible, which is conservatively matching the deadband used in SoilFluxPro. The CH~4~ data could not be fitted with the HM model because of the absence of curvature."

library(ggplot2)
library(ggpmisc)
library(viridis)



# comparison_fluxes_hm20 <- readRDS("comparison_fluxes_hm20ii.rds")

# flux_flag_count(comparison_fluxes_hm20)

gas_names <- c(
  `LI-7810_FCH4_DRY` = "a) LI-7810 CH4",
  `LI-7810_FCO2_DRY` = "b) LI-7810 CO2",
  `LI-7820_FN2O_DRY` = "c) LI-7820 N2O",
  `LI-870_FCO2_DRY` = "d) LI-870 CO2"
)

plot_sfp_hm <- comparison_fluxes_hm20 |>
    filter(
      method != "LI-7810_FCH4_DRY"
    ) |>
    ggplot(aes(f_flux, licor_flux, color = model)) +
    theme_bw() +
    scale_color_manual(values = c(
        Exponential = "#0c9f0b",
        Linear = "#bf28bd")
    ) +
    # theme(legend.position = c(0.87, 0.25)) +
    # scale_fill_viridis(discrete = TRUE) +
    geom_point() +
    geom_abline(slope = 1, intercept = 0) +
    stat_poly_line() +
    stat_correlation(use_label("cor.label", "R2", "n")) +
    # stat_poly_eq(use_label("R2", "n")) +
    facet_wrap(gas ~ ., scales = "free", nrow = 2, labeller = as_labeller(gas_names)) +
    labs(
        title = "Comparison of flux estimates between SoilFluxPro and Fluxible",
        x = "Fluxes calculated with Fluxible (exp_hm and linear model respectively)",
        y = "Fluxes as provided by LI-COR (calculated with SFP)",
        color = "Model"
    )
plot_sfp_hm
```

## CH~4~ further investigation for revision

```{r}
#| label: ch4-reinvestigation
#| echo: true
#| message: FALSE
#| warning: false

ch4_7810 <- readRDS("validation_data/ch4_7810.rds")

ch4_7810_fit <- flux_fitting(ch4_7810, start_cut = 20, fit_type = "linear")

ch4_7810_flags <- flux_quality(ch4_7810_fit, instr_error = 10, error = 100000) #just getting rid of the start error since we have different gases and units

# 3925 fluxes out of 16689 are flagged as zero, let's investigate that

# isolate the zero flags

zero_id <- ch4_7810_flags |>
    filter(f_quality_flag == "zero") |>
    distinct(f_fluxid) |>
    pull(f_fluxid)

ch4_7810_fit_a <- ch4_7810 |>
    filter(
        !(f_fluxid %in% zero_id)
        ) |>
    flux_fitting(start_cut = 20, fit_type = "linear")

ch4_7810_flags_a <- flux_quality(ch4_7810_fit_a, instr_error = 10, error = 100000)

ch4_7810_fit_b <- ch4_7810 |>
    filter(
        (f_fluxid %in% zero_id)
        ) |>
    flux_fitting(start_cut = 40, fit_type = "linear") # this has been investigated already and solved a big part of the issue

ch4_7810_flags_b <- flux_quality(ch4_7810_fit_b, instr_error = 10, error = 100000)

# still 2846 zero flags

zero_id_b <- ch4_7810_flags_b |>
    filter(f_quality_flag == "zero") |>
    distinct(f_fluxid) |>
    pull(f_fluxid)

# random plotting of 100 fluxes to see what is going on
random_id <- sample(zero_id_b, 100)

ch4_7810_flags_b |>
    filter(
        f_fluxid %in% random_id
    ) |>
    flux_plot(
        f_ylim_upper = 2100,
        f_ylim_lower = 2000,
        output = "longpdf",
        y_text_position = 2050,
        f_plotname = "random_100_ch4_zero_2"
    )

# those fluxes are very small and that affects the R2 

ch4_7810_flags_c <- flux_quality(ch4_7810_fit_b,
    instr_error = 10,
    error = 100000,
    rsquared_threshold = 0.01 # "deactivating" the R2 threshold
    )

```

We will redo the data processing for CH~4~ with an R² threshold of 0.01, which will stop flagging those very low fluxes as "zero".
It makes sense because (1) SFP does not do zero flagging, and (2) if a dataset contains only such low fluxes (as this is the case here) applying a high R² threshold might not be appropriate as it will discard or zero a large part of the fluxes.

```{r}
#| label: cleaning
#| include: false

unlink("validation_data", recursive = TRUE)
```

\newpage

# Example data

```{r}
#| label: example-import
#| message: true
#| warning: false
library(dataDownloader)

get_file(
    node = "srj24",
    file = "ex_data.zip"
)

unzip("ex_data.zip")

# saving space
file.remove("ex_data.zip")
```


